Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-25 draft.tex
several fixes

\newcommand{\profile}{\bm{v}}%(complete) profile
\newcommand{\pprofile}{{\bm{p}}}%partial profile
\newcommand{\w}{\bm{w}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\pw}{W}%our knowledge about the weights
\newcommand{\powersetz}[1]{\mathscr{P}^*(#1)}

@@

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full preference information is provided by the voters. 
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting voters information have been proposed.

In this paper we go one step further and we assume that both the voting rule and the voters preferences are partially specified.
Focusing on positional scoring rules, we assume that the chair, while  not able to give a precise definition of the rule, is capable of answering simple questions  requiring to pick a winner from a specific  example profile.
Moreover, the preferences of the voters are incrementally acquired by asking comparison queries.

We propose a method  for robust approximate winner determination in this setting with minimax regret. 
We then provide an interactive elicitation protocol based on minimax regret
and develop several query strategies that interleave questions to the chair and questions to the voters in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
\end{abstract}

@@

In this paper we consider that both the voters’ preferences and the social choice rule are partially specified.
We develop methods for computing the minimax-optimal alternative for scoring rules with partial voter preference information and partial information about the weights of the scoring rule.
While previous works have considered either partial information about the voters preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.

@@

%In this paper we propose an interactive adaptive protocol for eliciting both the voters preferences and the voting rule.
%At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).
The paper is organized as follows.
In Section \ref{sec:background} we provide the necessary background.
We introduce the minimax criterion, that selects a winner that minimizes the worst possible regret, in Section \ref{sec:mmr}.
In Section \ref{sec:elicit} we provide an interactive elicitation protocol based on minimal regret;  in Section \ref{sec:experiments} we present the empirical validation of our approach with simulations; and Section \ref{sec:conclusions} provides some final thoughts.

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We assume defined a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents (voters); each agent $j$ is associated to her “real” preference order $\pref_j$. The agents come from an infinite set $\N$ of potential voters.
The real preferences of the agents are elements of $\linors$, the linear orders (connected, transitive, asymmetric relations) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each voter, considering a subset of voters from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

@@

In this work we assume fixed, but unknown, a profile $\profile^*$ (and corresponding voters preferences $\pref_j^*$) and a weight vector $\w^*$, and we want to reason about partial preference information concerning those objects.
Our knowledge at a given time of the preference of voter $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each voter to a partial preference.

@@

An important class of scoring rule is the one composed of weights that are a convex sequence \cite{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as great as the difference between the weights of the second and third positions, etc.
%\begin{align} \forall r \in \{1,\ldots,m-2\} \;\; & w_r - w_{r+1} \geq w_{r+1}-w_{r+2}  \\
%\iff & w_r - 2 w_{r+1} + w_{r+2} \geq 0 \label{eq:convexity}
%\end{align}
\begin{equation} 
\label{eq:convexity}
\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\W$ to denote the set of convex weight vectors.

In general it can be difficult to set the weights in an appropriate way.
We assume that in addition of basic requirements (monotonicity and convexity), the chair (the person or the organization that is supervising the voting process) is able to specify additional preferences about how the social choice function should behave.
In this work we assume that preferences of the chair are encoded with linear constraints about the vector $\w$, relating the value of the weights of different positions.
We use $\pw \subseteq \W$ to denote the set of weight vectors compatible with the preferences expressed by the chair about the scoring vector, and $C_W$ to denote the set of linear constraints expressed by the chair.

@@

%The quality of an alternative can be quantified by considering the maximum regret with respect to an adversary that can choose the instantiation of both a complete profile (extending the known preferences of the agents) and of the scoring vectors (associated to the preferences of the chair).
In this paper, we consider a setting where both the voters' preferences and the preferences of the chair about the voting rule are incomplete.
Notice that some authors have considered possible and necessary winners assuming a partial profile  \citep{Xia2008} or assuming an incompletely specified scoring rule \citep{Viappiani2018};
however, we note that, in typical settings, there will be no necessary winner and too many possible winners.
In practice, however, it is often useful to declare a winner given partial information.

As a decision criterion to determine a winner, we propose to use minimax regret. 
Minimax regret \citep{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
\citet{Lu2011} have adopted minimax regret for winner determination in social choice with
the preferences of the voters that are only partially known, while the social choice function is predetermined and known.

In this work we consider the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
Using {\em maximum regret} to quantify the worst-case error, the alternative that minimize this error is selected as winner, providing us with a form of robust optimization.
Intuitively, the quality of a proposed alternative $a$ is how far from optimal $a$  could be in the worst case, given the current knowledge about the voting rule and about the voters' preferences.

@@

We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
	There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies \ref{eq:compl}.
\end{claim}
In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.

It follows that the rank of $x$ for voter $j$ in $\PMR(x,y; \pprofile, W)$ is $1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $1+\card{{\pprefinv_j}(y)}+\alpha$, where $\alpha = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\alpha = 0$ otherwise.

@@

%Let $U^+=\{ i \mid \ y \ppref_i x\}$ be the set of agents for which we know that $y$ is preferred to $x$ (positive contribution to pairwise regret), $U^-=\{ i \mid x \ppref_i y\}$ 
%be the set of agents for wich we  know that $x$ is preferred to $y$ (negative contribution) and 
%$U^?=\{ i \mid x \pinc_i y\}$ 
%%$U^?=\{ i \mid x \nsucc_{i} y \wedge y \nsucc_{i} x \}$ 
%the remaining case, where the preference between $x$ and $y$ is not known.
%Consider how the adversary, whose goal is to maximize $\PMR$ between $x$ and $y$, will complete the partial orders:
%\begin{itemize}
% \item $j \in U^+$: If we know that agent $j$ prefers $y$ to $x$ then the contribution to $\PMR$ is positive and adversary will complete the partial order of agent $j$ by placing as many alternatives as possible between $y$ and $x$ (because of monotonicity of the weights).
% The way the partial order is completed does not depend on the actual values of the weights.
%
% \item $j \in U^?$: If we do not know whether $x$ or $y$ is preferred by agent $j$ then the adversary will place $y$ before $x$ (since the goal is to maximize the difference in score between $y$ and $x$, and weights are monotone) in the linear order and this case reduce to the first one; therefore also in this case the completion of the partial order can be done independently of the choice of $w$.
%
% \item $j \in U^-$: If we know that agent $j$ prefers $x$ to $y$ then (as the contribution of this agent to $\PMR$ is negative) the adversary will place as few alternatives as possible between $x$ and $y$.
%
%The only issue is with respect to the  alternatives that are incomparable with $x$ and $y$ according to $\ppref_i$: should they be placed better than $x$ or worse than $y$ ?
%If the scoring vector is known to be convex,  than these alternatives will be placed better than $x$ in the completion; otherwise the placements of these depend on the weights $\w$.
%\end{itemize}

\subsubsection{Minimax regret computation}

%Since weights are assumed to be convex, the adversary will place all alternatives $c$ that are such that $c \pinc_{j} y \wedge c \pinc_{j} x$ as better than $x$.
%We now formalize the reasoning about the best (with respect to maximizing $\PMR$) completion with the following definition.
%\begin{definition}
%Given the partial profile $\pprofile$, define $\hat{\profile}$ as the completition of $\pprofile$ such that:
%\begin{itemize}
%\item For each $j \in U^{+}$, set   $\hat{\profile}$ as follows, for all $c \in A$:\\
%if $c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%if  $c \pinc_{j} y$ we set $y \succ_{j}^{\hat{\profile}} c$.
%\item For each $j \in U^{?}$, set  $\hat{\profile}$ as follows:\\
%$y \succ_{j}^{\hat{\profile}} x$;\\
%for all $c \in A$, if   $c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%for all $c \in A$, if   $c \pinc_{j} y$ we set $y \succ_{j}^{\hat{\profile}} c$.
%\item For each $j \in U^{-}$,  set  $\hat{\profile}$ as follows, for all $c \in A$:\\
%if $c \pinc_{j} x \wedge c \succ^{p}_{j} y$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%if $c \pinc_{j} y \wedge  x \succ^{p}_{j} c$ we set $y \succ_{j}^{\hat{\profile}} c$;\\
%if $c \pinc_{j} y \wedge c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$.
%\end{itemize}
%\end{definition}

@@

Let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
\begin{prop}
%Assume convex weights.
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W) =  
\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
=\max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} =
\max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
\end{align}
where the coefficient $ (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x})$ is the number of times $y$ is at rank $r$ minus the number of times $x$ is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{prop}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& C_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables. We leave $w_{1}$ and $w_{m}$ in the linear program just for clarity of presentation.

@@

The remainder of this Section is structured as follows.
First of all, we discuss the different types of questions that can be asked to the voters and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

\paragraph{Question types}
We distinguish between questions asked to the voters and questions asked to the chair.
As {\em questions asked to the voters}, it is natural to consider comparison queries asking to compare two alternatives.
%Another common type of queries are {\em top-k}, asking to each voter her $k$ most preferred alternatives.
The effect of a response to a question asked to a voter is to augment our knowledge about the voters rankings, thus augmenting the partial profile $\pprofile$. 
If voter $j$ answers a comparison query stating that alternative $a$ is preferred to $b$, then the partial order $\ppref_j$ is augmented with $a \ppref_j b$ and by transitive closure.

@@

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ voters. Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each voter in the resulting profile has given each rank once. And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$. For if the chair picks a different alternative $c$, it only tells us that the chair prefers $c$ to both $a$ and $b$, but it generally does not answer the question we are interested in. Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of voters.

First build a temporary profile $P$ of $p+q$ voters with $a$ and $b$ ranked as just described. Complete it with $m-2$ alternatives ranked in arbitrary orders so that the resulting profile has complete strict rankings for each of the $p+q$ voter. Now we will make the other alternatives as bad as desired by adding voters to $P$ that appreciate $a$ and $b$ more than the other alternatives, thus building a profile $P'$ in which, whatever the weights, $c$ may not have a better score than $a$. Observe that if we add $\delta$ voters (with $\delta$ a natural number) that put $a$ at first rank and $b$ at second rank, and $\delta$ voters that put $b$ at first rank and $a$ at second rank, we do not change the difference of the scores of $a$ and $b$, and thus, observing the chair choosing $a$ or $b$ still answers the question. Choose arbitrarily a matching of the remaining $m-2$ alternatives to the ranks $3$ to $m$, say, $c \mapsto 3$, $d \mapsto 4$, and so on. The first $\delta$ supplementary voters, who rank $a$ first and $b$ second, rank the other alternatives as the just defined mapping suggests. The other $\delta$ supplementary voters, who rank $b$ first and $a$ second, use the opposite ranking for the remaining alternatives, say, $c \mapsto m$, $d \mapsto m-1$, and so on. We can now prove that, for $\delta = p+q$, and whatever the weight vector $w \in \W$, no alternative, except possibly $b$, has a better score than $a$. (In fact, as the construction will exhibit, they are all worst than both $a$ and $b$, but the weaker fact is enough for our claim.)

>>>>

\newcommand{\profile}{\bm{v}}%(complete) profile
\newcommand{\pprofile}{{\bm{p}}}%partial profile
\newcommand{\w}{\bm{w}}
\newcommand{\W}{\mathcal{W}}
\newcommand{\Co}{\mathcal{C}}
\newcommand{\pw}{W}%our knowledge about the weights
\newcommand{\powersetz}[1]{\mathscr{P}^*(#1)}

@@

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full preference information is provided.% by the voters. 
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting preferences have been proposed.

In this paper we go one step further and we assume that both the voting rule and the agents preferences are partially specified.
Focusing on positional scoring rules, we assume that the chair, while  not able to give a precise definition of the rule, is capable of answering simple questions  requiring to pick a winner from a specific  example profile.
Moreover, the preferences of the agents are incrementally acquired by asking comparison queries.

We propose a method  for robust approximate winner determination in this setting with minimax regret. 
We then provide an interactive elicitation protocol based on minimax regret
and develop several query strategies that interleave questions to the chair and questions to the agents in order to acquire the most relevant information in order to quickly converge to optimal or a near-optimal alternative.
\end{abstract}

@@

In this paper we consider that both the agents’ preferences and the social choice rule are partially specified.
We develop methods for computing the minimax-optimal alternative for scoring rules with partial agent preference information and partial information about the weights of the scoring rule.
While previous works have considered either partial information about the agents preferences or a partially specified aggregation method, we do not know of any work considering both sources of uncertainty at the same time.

@@

%In this paper we propose an interactive adaptive protocol for eliciting both the agents preferences and the voting rule.
%At each step we need to decide whether we want to ask a question to the committee or to one of the agents (and to which agent in particular).
The paper is organized as follows.
In Section \ref{sec:background} we provide the necessary background.
We introduce the minimax criterion, that selects a winner that minimizes the worst possible regret, in Section \ref{sec:mmr}.
In Section \ref{sec:elicit} we provide an interactive elicitation protocol based on minimal regret;  in Section \ref{sec:experiments} we present the empirical validation of our approach with simulations; and Section \ref{sec:conclusions} provides some final thoughts.

\section{Social choice with partial information}
\label{sec:background}
We now introduce some basic concepts.
We assume defined a set $A$ of $m$ alternatives (products, restaurants, movies, public projects, job candidates, etc.) and a set $\set{1, …, n}$ of agents (voters); each agent $j$ is associated to her “real” preference order $\pref_j$. The agents come from an infinite set $\N$ of potential agents.
The real preferences of the agents are elements of $\linors$, the linear orders (connected, transitive, asymmetric relations) over the alternatives.
Following the social choice nomenclature, we call {\em profile} the association of a preference to each agent, considering a subset of agents from the set $\N$, and denote a profile by $(\pref_1,\ldots,\pref_n)$.
A profile is equivalently represented by $\profile=(v_1,\ldots,v_n)$ where $v_j(i) \in \set{1, \ldots, m}$ denotes the rank (position) of alternative $i$ in the preference order $\pref_j$.

@@

In this work we assume fixed, but unknown, a profile $\profile^*$ (and corresponding agents preferences $\pref_j^*$) and a weight vector $\w^*$, and we want to reason about partial preference information concerning those objects.
Our knowledge at a given time of the preference of voter $j$ is encoded by a partial order over the alternatives, thus a transitive and asymmetric binary relation, denoted by $\ppref_j$. 
In this work we assume that preference information is truthful, i.e. $a \ppref_j b ⇒ a \pref_j^* b$.
We use $\pinc$ to denote incomparability, that is $a \pinc_{j} b$ iff $a \nppref_j b \wedge b \nppref_j a$.
An incomplete profile $\pprofile = (\ppref_1, \ldots, \ppref_n)$ maps each voter to a partial preference.

@@

An important class of scoring rule is the one composed of weights that are a convex sequence \citep{Stein1994,Llamazares2016}, meaning that the difference between the weight of the first position and the weight of the second position is at least as great as the difference between the weights of the second and third positions, etc.
%\begin{align} \forall r \in \{1,\ldots,m-2\} \;\; & w_r - w_{r+1} \geq w_{r+1}-w_{r+2}  \\
%\iff & w_r - 2 w_{r+1} + w_{r+2} \geq 0 \label{eq:convexity}
%\end{align}
\begin{equation} 
\label{eq:convexity}
\forall r \in \{1,\ldots,m-2\}: w_r - w_{r+1} \geq w_{r+1}-w_{r+2}.
\end{equation}
The constraint above is often used when aggregating rankings in sport competitions.
We use $\W$ to denote the set of convex weight vectors.

In general it can be difficult to set the weights in an appropriate way.
We assume that in addition of basic requirements (monotonicity and convexity), the chair (the person or the organization that is supervising the voting process) is able to specify additional preferences about how the social choice function should behave.
In this work we assume that preferences of the chair are encoded with linear constraints about the vector $\w$, relating the value of the weights of different positions.
We use $\pw \subseteq \W$ to denote the set of weight vectors compatible with the preferences expressed by the chair about the scoring vector, and $\Co_W$ to denote the set of linear constraints expressed by the chair.

@@

%The quality of an alternative can be quantified by considering the maximum regret with respect to an adversary that can choose the instantiation of both a complete profile (extending the known preferences of the agents) and of the scoring vectors (associated to the preferences of the chair).
In this paper, we consider a setting where both the agents' preferences and the preferences of the chair about the voting rule are incomplete.
Notice that some authors have considered possible and necessary winners assuming a partial profile  \citep{Xia2008} or assuming an incompletely specified scoring rule \citep{Viappiani2018};
however, we note that, in typical settings, there will be no necessary winner and too many possible winners.
In practice, however, it is often useful to declare a winner given partial information.

As a decision criterion to determine a winner, we propose to use minimax regret. 
Minimax regret \citep{Savage1954} is a decision criterion that has been used for robust optimization under data uncertainty \citep{Kouvelis1997} as well as in decision-making with uncertain utility values \citep{Salo2001,Boutilier2006}.
\citet{Lu2011} have adopted minimax regret for winner determination in social choice with
the preferences of the agents that are only partially known, while the social choice function is predetermined and known.

In this work we consider the simultaneous presence of uncertainty in the agents' preferences and uncertainty in the weights.
Using {\em maximum regret} to quantify the worst-case error, the alternative that minimize this error is selected as winner, providing us with a form of robust optimization.
Intuitively, the quality of a proposed alternative $a$ is how far from optimal $a$  could be in the worst case, given the current knowledge about the voting rule and about the agents' preferences.

@@

We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
	There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies \ref{eq:compl}.
\end{claim}
In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
It follows that the rank of $x$ for voter $j$ in $\PMR(x,y; \pprofile, W)$ is $1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $1+\card{{\pprefinv_j}(y)}+\alpha$, where $\alpha = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\alpha = 0$ otherwise.

@@

%Let $U^+=\{ i \mid \ y \ppref_i x\}$ be the set of agents for which we know that $y$ is preferred to $x$ (positive contribution to pairwise regret), $U^-=\{ i \mid x \ppref_i y\}$ 
%be the set of agents for wich we  know that $x$ is preferred to $y$ (negative contribution) and 
%$U^?=\{ i \mid x \pinc_i y\}$ 
%%$U^?=\{ i \mid x \nsucc_{i} y \wedge y \nsucc_{i} x \}$ 
%the remaining case, where the preference between $x$ and $y$ is not known.
%Consider how the adversary, whose goal is to maximize $\PMR$ between $x$ and $y$, will complete the partial orders:
%\begin{itemize}
% \item $j \in U^+$: If we know that agent $j$ prefers $y$ to $x$ then the contribution to $\PMR$ is positive and adversary will complete the partial order of agent $j$ by placing as many alternatives as possible between $y$ and $x$ (because of monotonicity of the weights).
% The way the partial order is completed does not depend on the actual values of the weights.
%
% \item $j \in U^?$: If we do not know whether $x$ or $y$ is preferred by agent $j$ then the adversary will place $y$ before $x$ (since the goal is to maximize the difference in score between $y$ and $x$, and weights are monotone) in the linear order and this case reduce to the first one; therefore also in this case the completion of the partial order can be done independently of the choice of $w$.
%
% \item $j \in U^-$: If we know that agent $j$ prefers $x$ to $y$ then (as the contribution of this agent to $\PMR$ is negative) the adversary will place as few alternatives as possible between $x$ and $y$.
%
%The only issue is with respect to the  alternatives that are incomparable with $x$ and $y$ according to $\ppref_i$: should they be placed better than $x$ or worse than $y$ ?
%If the scoring vector is known to be convex,  than these alternatives will be placed better than $x$ in the completion; otherwise the placements of these depend on the weights $\w$.
%\end{itemize}
%\subsubsection{Minimax regret computation}
%Since weights are assumed to be convex, the adversary will place all alternatives $c$ that are such that $c \pinc_{j} y \wedge c \pinc_{j} x$ as better than $x$.
%We now formalize the reasoning about the best (with respect to maximizing $\PMR$) completion with the following definition.
%\begin{definition}
%Given the partial profile $\pprofile$, define $\hat{\profile}$ as the completition of $\pprofile$ such that:
%\begin{itemize}
%\item For each $j \in U^{+}$, set   $\hat{\profile}$ as follows, for all $c \in A$:\\
%if $c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%if  $c \pinc_{j} y$ we set $y \succ_{j}^{\hat{\profile}} c$.
%\item For each $j \in U^{?}$, set  $\hat{\profile}$ as follows:\\
%$y \succ_{j}^{\hat{\profile}} x$;\\
%for all $c \in A$, if   $c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%for all $c \in A$, if   $c \pinc_{j} y$ we set $y \succ_{j}^{\hat{\profile}} c$.
%\item For each $j \in U^{-}$,  set  $\hat{\profile}$ as follows, for all $c \in A$:\\
%if $c \pinc_{j} x \wedge c \succ^{p}_{j} y$ we set $c \succ_{j}^{\hat{\profile}} x$;\\
%if $c \pinc_{j} y \wedge  x \succ^{p}_{j} c$ we set $y \succ_{j}^{\hat{\profile}} c$;\\
%if $c \pinc_{j} y \wedge c \pinc_{j} x$ we set $c \succ_{j}^{\hat{\profile}} x$.
%\end{itemize}
%\end{definition}

@@

Let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
\begin{claim}
%Assume convex weights.
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W) =  
\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
=\max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} =
\max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i 
\end{align}
where the coefficient $ (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x})$ is the number of times $y$ is at rank $r$ minus the number of times $x$ is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{claim}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
The pairwise max regret $\PMR(x,y; \pprofile,W)$ is now obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$. 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& \Co_W.
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables. We leave $w_{1}$ and $w_{m}$ in the linear program just for clarity of presentation.

@@

The remainder of this Section is structured as follows.
First of all, we discuss the different types of questions that can be asked to the agents and to the chair, and the way responses are handled.
Then, we describe different strategies to determine informative queries to ask next, with the goal of reducing $\MMR(\pprofile,W)$ quickly.

\paragraph{Question types}
We distinguish between questions asked to the agents and questions asked to the chair.
As {\em questions asked to the agents}, it is natural to consider comparison queries asking to compare two alternatives.
%Another common type of queries are {\em top-k}, asking to each voter her $k$ most preferred alternatives.
The effect of a response to a question asked to a voter is to augment our knowledge about the agents rankings, thus augmenting the partial profile $\pprofile$. 
If voter $j$ answers a comparison query stating that alternative $a$ is preferred to $b$, then the partial order $\ppref_j$ is augmented with $a \ppref_j b$ and by transitive closure.

@@

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ agents. Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each voter in the resulting profile has given each rank once. And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$. For if the chair picks a different alternative $c$, it only tells us that the chair prefers $c$ to both $a$ and $b$, but it generally does not answer the question we are interested in. Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of agents.

First build a temporary profile $P$ of $p+q$ agents with $a$ and $b$ ranked as just described. Complete it with $m-2$ alternatives ranked in arbitrary orders so that the resulting profile has complete strict rankings for each of the $p+q$ voter. Now we will make the other alternatives as bad as desired by adding agents to $P$ that appreciate $a$ and $b$ more than the other alternatives, thus building a profile $P'$ in which, whatever the weights, $c$ may not have a better score than $a$. Observe that if we add $\delta$ agents (with $\delta$ a natural number) that put $a$ at first rank and $b$ at second rank, and $\delta$ agents that put $b$ at first rank and $a$ at second rank, we do not change the difference of the scores of $a$ and $b$, and thus, observing the chair choosing $a$ or $b$ still answers the question. Choose arbitrarily a matching of the remaining $m-2$ alternatives to the ranks $3$ to $m$, say, $c \mapsto 3$, $d \mapsto 4$, and so on. The first $\delta$ supplementary agents, who rank $a$ first and $b$ second, rank the other alternatives as the just defined mapping suggests. The other $\delta$ supplementary agents, who rank $b$ first and $a$ second, use the opposite ranking for the remaining alternatives, say, $c \mapsto m$, $d \mapsto m-1$, and so on. We can now prove that, for $\delta = p+q$, and whatever the weight vector $w \in \W$, no alternative, except possibly $b$, has a better score than $a$. (In fact, as the construction will exhibit, they are all worst than both $a$ and $b$, but the weaker fact is enough for our claim.)

Pick any alternative $c$ that is not $a$ or $b$. 
%Let $c_t$ denote the number of times $c$ reaches rank $t$ or better in $P$. Note that $c_m = p+q$ and $c_t ≤ p + q, 1 ≤ t ≤ m$. Define $a_t$ similarly, for the alternative $a$. By construction, $a_r = q$, $a_{r+2} = q+p = a_m$. 
Define $c'_t$ as the number of times $c$ gets rank $t$ or better in $P'$, and define $a'_t$ similarly. Now $a'_1 ≥ \delta, a'_t ≥ 2\delta \text{ for } 2 ≤ t < m, a'_{r+2} = q+p+2\delta = a'_m$. To obtain upper bounds for $c'_t$, assume $c$ is ranked first by the $p+q$ agents voting in profile $P$. By our construction, the two times $\delta$ new agents of $P'$ give to $c$ the ranks $(3, m)$, or $(4, m-1)$, … Observe that the score of $c$ is maximal for attribution $(3, m)$, by convexity of the weights. Therefore, in the best case for $c$, $c'_1 ≤ p + q, c'_2 ≤ p+q, c'_t ≤ p+q+\delta \text{ for } 3 ≤ t < m, c'_m ≤ p+q+2\delta$. Observe that $a'_t ≥ c'_t, 1 ≤ t ≤ m$ (with strict inequality for $t=2$). Because weights are non increasing, this guarantees, with $\delta = p + q$, that the score of $a$ is not lower than the one of $c$ (it is even strictly higher, thanks to the strict inequality). This concludes the proof of the following claim.
